---
title: "FLORA Converter - PubPeer Comments"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
```

## Load Converter

```{r load-converter}
source("flora_converter.R")
```

## Helper Functions

```{r helper-functions}
shorten_apa <- function(ref) {
  if (is.na(ref) || nchar(trimws(ref)) == 0) {
    return(ref)
  }

  year_match <- regmatches(ref, regexpr("\\(\\d{4}\\)", ref))
  if (length(year_match) == 0) {
    return(ref)
  }

  year <- year_match[1]

  author_str <- trimws(strsplit(ref, year, fixed = TRUE)[[1]][1])
  first_author_family <- trimws(strsplit(author_str, ",")[[1]][1])
  comma_count <- lengths(regmatches(author_str, gregexpr(",", author_str)))
  rest <- paste0(year, trimws(strsplit(ref, year, fixed = TRUE)[[1]][2]))

  if (comma_count >= 3) {
    paste0(first_author_family, " et al. ", rest)
  } else {
    paste0(author_str, " ", rest)
  }
}
```

## Load And Clean Data

```{r load-clean-data}
data <- csv_to_flora_json("flora.csv")
df <- flora_json_to_df(data)

df <- df |>
  mutate(
    across(
      where(is.character),
      \(x) iconv(x, from = "latin1", to = "UTF-8", sub = " ")
    )
  ) |>
  mutate(across(where(is.character), \(x) str_replace_all(x, "[^\x01-\x7F]", "-"))) |>
  mutate(across(where(is.character), \(x) str_replace_all(x, "-{2,}", "-"))) |>
  mutate(across(where(is.character), \(x) str_replace_all(x, "\\s+", " "))) |>
  mutate(across(where(is.character), \(x) trimws(x)))

df$apa_ref_r_sh <- sapply(df$apa_ref_r, shorten_apa)

rand_doi <- sample(df$doi_o, 200)
df_rand50 <- df |>
  filter(doi_o %in% rand_doi)
```

## Generate PubPeer Messages

The message generation workflow:

1. Convert JSON back to a data frame.
2. Keep records with a non-missing `doi_o` and a value in either `doi_r` or `url_r`.
3. Create a CSV file for output.
4. Generate one row per unique `doi_o`, including the reference, outcome, quote, quote source, and message. If there are multiple `doi_r` values, include all of them in the message.

```{r generate-messages-function}
generate_flora_messages <- function(df) {
  messages <- list()

  for (doi in unique(df$doi_o)) {
    rows <- df[df$doi_o == doi, ]
    rows$apa_ref_r_sh <- sapply(rows$apa_ref_r, shorten_apa)
    n <- nrow(rows)

    n_replications <- sum(tolower(rows$type) == "replication", na.rm = TRUE)
    n_reproductions <- sum(tolower(rows$type) == "reproduction", na.rm = TRUE)

    attempt_parts <- c(
      if (n_replications > 0) {
        paste0(
          n_replications,
          " replication attempt",
          if (n_replications != 1) "s" else ""
        )
      },
      if (n_reproductions > 0) {
        paste0(
          n_reproductions,
          " reproduction attempt",
          if (n_reproductions != 1) "s" else ""
        )
      }
    )

    header <- paste0(
      "**",
      paste(attempt_parts, collapse = " and "),
      " of this paper ",
      if ((n_replications + n_reproductions) != 1) "have" else "has",
      " been recorded in the FORRT Library of Reproduction and ",
      "Replication Attempts (FLoRA).",
      "**"
    )

    rep_counter <- 0
    repro_counter <- 0

    entries <- vapply(seq_len(n), function(i) {
      r <- rows[i, ]

      if (n == 1) {
        label <- ""
      } else if (tolower(r$type) == "replication") {
        rep_counter <<- rep_counter + 1
        label <- paste0("**Replication ", rep_counter, "**\n")
      } else {
        repro_counter <<- repro_counter + 1
        label <- paste0("**Reproduction ", repro_counter, "**\n")
      }

      quote_source_line <- if (!is.na(r$outcome_quote_source)) {
        paste0("\n> **Quote source:** \"", r$outcome_quote_source, "\"")
      } else {
        ""
      }

      oa_link <- if (!is.na(r$oa_url_r) && nchar(trimws(r$oa_url_r)) > 0) {
        paste0(" [Open Access](", r$oa_url_r, ")")
      } else {
        ""
      }

      paste0(
        label,
        "\n> ", r$apa_ref_r_sh, oa_link, "\n",
        "> \n",
        "> **The authors reported the replication as:** \"", r$outcome, "\"\n",
        "> \n",
        "> **Quote:** ", r$outcome_quote,
        quote_source_line
      )
    }, character(1))

    footer <- paste0(
      "Manual inspection may help to assess what led to the outcome described. ",
      "Future research can learn from replications and build on them. More ",
      "information about why and how we submitted this automated comment can be ",
      "found [here](https://github.com/forrtproject/flora-pubpeer/blob/main/README.md)\n\n",
      "FLoRA Notify Team"
    )

    messages[[doi]] <- paste(c(header, entries, footer), collapse = "\n\n\n")
  }

  messages
}
```

## Create Message CSV

```{r create-message-csv}
msgs <- generate_flora_messages(df_rand50)

msgs_df <- data.frame(
  doi_o = names(msgs),
  message = unlist(msgs),
  row.names = NULL,
  stringsAsFactors = FALSE
) |>
  mutate(
    message = iconv(message, from = "UTF-8", to = "UTF-8", sub = " "),
    message = str_replace_all(message, "[^\x01-\x7F]", "-"),
    message = str_replace_all(message, "-{2,}", "-"),
    message = str_replace_all(message, "\\[Open Access\\]", "\U0001F513 [Open Access]")
  )

msgs_df |>
  filter(doi_o == "10.1016/j.obhdp.2006.10.001")

readr::write_csv(msgs_df, "msgs.csv")
```
