---
title: "FLORA Converter — Notebook Test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)

```

## Load the converter

```{r}
source("flora_converter.R")
```

# Shorten references length

```{r}
shorten_apa <- function(ref) {
  if (is.na(ref) || nchar(trimws(ref)) == 0) return(ref)
  
  # Extract year
  year_match <- regmatches(ref, regexpr("\\(\\d{4}\\)", ref))
  if (length(year_match) == 0) return(ref)
  year <- year_match[1]
  
  # Extract author string — everything before the year
  author_str <- trimws(sub(year, "", ref, fixed = TRUE))
  author_str <- trimws(sub(paste0(sub(year, "", ref, fixed = TRUE) |> 
                                    strsplit(year, fixed = TRUE) |> 
                                    unlist() |> tail(1)), "", author_str, fixed = TRUE))
  author_str <- trimws(strsplit(ref, year, fixed = TRUE)[[1]][1])
  
  # Get first author family name
  first_author_family <- trimws(strsplit(author_str, ",")[[1]][1])
  
  # Count commas to determine number of authors
  comma_count <- lengths(regmatches(author_str, gregexpr(",", author_str)))
  
  # Build rest of reference from year onward
  rest <- paste0(year, trimws(strsplit(ref, year, fixed = TRUE)[[1]][2]))
  
  if (comma_count >= 3) {
    paste0(first_author_family, " et al. ", rest)
  } else {
    paste0(author_str, " ", rest)
  }
}


```



# load data convert
```{r}

data <- csv_to_flora_json("flora.csv")
df   <- flora_json_to_df(data)

View(flora_json_to_df)

# Fix encoding immediately after loading
df <- df |>
  mutate(across(where(is.character), \(x) iconv(x, from = "latin1", to = "UTF-8", sub = " "))) |>
  mutate(across(where(is.character), \(x) str_replace_all(x, "[^\x01-\x7F]", "-"))) |>
  mutate(across(where(is.character), \(x) str_replace_all(x, "-{2,}", "-"))) |>  # ← add this line
  mutate(across(where(is.character), \(x) str_replace_all(x, "\\s+", " "))) |>
  mutate(across(where(is.character), \(x) trimws(x)))


# Create shortened APA reference
df$apa_ref_r_sh <- sapply(df$apa_ref_r, shorten_apa)


# Select 50 random rows
rand_doi <- sample(df$doi_o, 200)
df_rand50 <- df |> filter(doi_o %in% rand_doi)



```



# process
1. convert json back to data frame
2. clean up so that only non na doi_o and value in either doi_r or url_r
3. new csv file for output
4. for each unique doi_o generate row with reference, outcome, outcome quote and outcome source columns, column with message. If multiple doi_r should include all in message.

```{r}
generate_flora_messages <- function(df) {
  messages <- list()

  for (doi in unique(df$doi_o)) {
    rows <- df[df$doi_o == doi, ]
    rows$apa_ref_r_sh <- sapply(rows$apa_ref_r, shorten_apa)
    n    <- nrow(rows)

      n_replications  <- sum(tolower(rows$type) == "replication", na.rm = TRUE)
      n_reproductions <- sum(tolower(rows$type) == "reproduction", na.rm = TRUE)
      

      attempt_parts <- c(
        if (n_replications > 0) paste0(n_replications, " replication attempt", if (n_replications != 1) "s" else ""),
        if (n_reproductions > 0) paste0(n_reproductions, " reproduction attempt", if (n_reproductions != 1) "s" else "")
      )
      
      header <- paste0(
        "**",
        paste(attempt_parts, collapse = " and "),
        " of this paper ",
        if ((n_replications + n_reproductions) != 1) "have" else "has",
        " been recorded in the FORRT Library of Reproduction and Replication Attempts (FLoRA).",
        "**"
      )

   rep_counter <- 0
repro_counter <- 0

entries <- vapply(seq_len(n), function(i) {
            r <- rows[i, ]
            
            # Build the label
            if (n == 1) {
              label <- ""
            } else if (tolower(r$type) == "replication") {
              rep_counter <<- rep_counter + 1
              label <- paste0("**Replication ", rep_counter, "**\n")
            } else {
              repro_counter <<- repro_counter + 1
              label <- paste0("**Reproduction ", repro_counter, "**\n")
            }
            
            quote_source_line <- if (!is.na(r$outcome_quote_source)) paste0("\n> **Quote source:** \"",
                                                                              r$outcome_quote_source, "\"") else ""
            
            oa_link <- if (!is.na(r$oa_url_r) & nchar(trimws(r$oa_url_r)) > 0) {
                        paste0(" [🔓](", r$oa_url_r, ")")
                      } else {
                        ""
                      }
                      
                      paste0(
                        label,
                        "\n> ", r$apa_ref_r_sh, oa_link, "\n",
              "> \n",
              "> **The authors reported the replication as:** \"", r$outcome, "\"\n",
              "> \n",
              "> **Quote:** ", r$outcome_quote,
              quote_source_line
            )
          }, character(1))
   
   
    

    footer <- paste0(
      "Manual inspection may help to assess what led to the outcome described. Future research can learn from replications and build on them. More information about why and how we submitted this automated comment was submitted can be found [here](https://github.com/forrtproject/flora-pubpeer/blob/main/README.md)\n\n",
      "FLoRA Notify Team"
    )
    messages[[doi]] <- paste(c(header, entries, footer), collapse = "\n\n\n")
  }

  messages
}


```

# create message
```{r}
msgs_df <- data.frame(
  doi_o   = names(msgs),
  message = unlist(msgs),
  row.names = NULL,
  stringsAsFactors = FALSE
) |>
  mutate(
    message = iconv(message, from = "UTF-8", to = "UTF-8", sub = " "),
    message = str_replace_all(message, "[^\x01-\x7F]", "-"),
    message = str_replace_all(message, "-{2,}", "-")
  )




msgs_df %>% filter(doi_o=="10.1016/j.obhdp.2006.10.001")
readr::write_csv(msgs_df, "msgs.csv")


```
:::